% !TEX root =  main.tex
I would like to thank...
\begin{description}
    \item[My family,] for making sure I don't have to think about more things than I have to throughout my learning years, even in college. Thank you also for always providing more than what I need, and sometimes even more than what I want. I couldn't have afforded to be so carefree without you.
    \item[Dr. Johnrob Bantang,] my adviser, for your patience, understanding, and guidance, for often treating us with food, and for the corniest and nerdiest jokes.
    \item[The Complexity Science Group.] Reinier: for always being someone we can rely on and relate with. Laya and Sara: for making sure I felt at home with the group, especially when I first joined. Jeruine: for all the times we were lost together, both in class and in research, but figured things out one way or another.
    \item[The Flahtotoys,] who are always there to support me with whatever problems I have, to share in my nerdy enthusiasm, and to go on so many adventures with. Special thanks to Hanzel who would come to my rescue even at such short notice. Life would not be as fun or tolerable if I hadn't met you all.
    \item[Julia,] for believing in me and for appreciating me simply for who I am, flaws and all. Thank you for doing your very best to always be there for me, for always being patient and understanding with me, and for always being proud of me no matter what mistake I make.
    \item[Janella,] for always matching my energy when I talk about literally anything, and for being someone I can confide in without fear of judgement.
    \item[The Hexagon,] Matt and Darwin, for being the constants that you are since we met and for being people I can talk to about my problem(set)s, with topics ranging from electrodynamics and quantum mechanics to where we'll be eating lunch. You have helped me stay sane and have been constant reminders that there's much joy in being nerds together, even if what we end up talking about isn't remotely nerdy.
    \item[Ms. Buenaventura,] for being someone I can talk to about my problems, and for being someone who can give me advice on how to deal with them.
    \item[The 401 tambays,] Belle, Eli, Edneil, and Mika, for always being there to keep me company when I don't feel like going home for the day yet, and for inviting me to your occasional lakwatsas.
    \item[The Stephenian Scout Group,] for always welcoming me back to a place where I can do what I enjoy outside the academe. Marc: for being someone I can turn to for knowledge and wisdom about the so many interests we share. Sid, Francynne, Chester, Kim, and Sam: for always being supportive and being g for the most random galas.
    \item[The Cakes,] for being people I can talk about the latest meta and math for our favorite gacha games, and for sharing the struggles of being college students even if we're all in different schools.
    \item[Monica,] for being unfathomably patient and supportive when I needed someone to be just that.
\end{description}

I can't list them all, but none of what any of you have done goes unappreciated. From the bottom of my heart, thank you.
